---
title: "Bootstrap Resampling Method"
output: html_document
---

<br> 

Two step process. First - know your parameter of interest. Second - use the boot function. 

<br> 

```{r}
library('boot') # for boot()
library('ISLR') # for portfolio data
```

<br>

Create function - 

```{r}
alpha.fn <- function(data, index) {
  X=data$X[index]
  Y=data$Y[index]
  return((var(Y)-cov(X,Y))/(var(X)+var(Y)-2*cov(X,Y)))
}
```

<br> 

Test function - 

```{r}
alpha.fn(Portfolio, 1:100)
```

<br> 

sampling with replacement is basically creating a bootstrap

```{r}
set.seed(1)
alpha.fn(Portfolio, sample(100, 100, replace = T))
```

<br> 

We can do this many times and record all of the output values. the boot function automates this process.

```{r}
boot(Portfolio, alpha.fn, 1000)
```

<br> 

This returns the original alpha calculation based on the real dataset, and a standard error estimate of this calculation based on the 1000 bootstraps. 

<br>

## Estimating Accuracy of a Linear Model 

<br> 

Just how good are $\beta_0$ and $\beta_1$ anyways?

```{r}
# create a bootstrap function for coefficients of linear model 
boot.fn <- function(data, index) {
  return(coef(lm(mpg ~ horsepower, data = data, subset = index)))
}

# test function
boot.fn(Auto, 1:392)
```

<br> 

The above is just returning the coeffificents of the linear model based on the entire auto dataset, observations 1 - 392. But we can use this function to return bootstrap coefficients. 

<br> 

```{r}
set.seed(1)
boot.fn(Auto, sample(392, 392, replace = T))
boot.fn(Auto, sample(392, 392, replace = T))
```

<br> 

Now let's bring out the big guns - the boot() function. 

```{r}
boot(Auto, boot.fn, 1000)
```

<br> 

This bootstrap estimate for the standard error of the intercept is 0.86, and the standard error of $\beta_1$, the slope, is 0.007. 

