---
title: "Cross Validation"
output: html_document
---


## Simple Cross Validation

The basic idea here is just to split your dataset into two parts, one for training, one for testing. This is called cross validation. 

<br> 

Load library, set seed, declare training sample. 

```{r}
library('ISLR')

set.seed(1)

train <- sample(392, 196)
```

<br> 

This is what we're attempting to model: 

```{r}
plot(Auto$horsepower, Auto$mpg)
```

<br>

Create a simple linear regression based on the auto dataset

```{r}
model_1 <- lm(mpg ~ horsepower, data = Auto, subset = train)
```


<br> 

Create predictions for the test dataset using the model, then calculate the MSE of these predictions.

```{r}
pred_1 <- predict(model_1, Auto)[-train]

mean((Auto$mpg[-train] - pred_1)^2)
```

<br> 

The MSE of a simple linear model is 26.14. Now lets use a quadratic and cubic model, then cross validate to compare resulting MSE's. 

```{r}
model_2 <- lm(mpg ~ poly(horsepower, 2), data = Auto, subset = train)
model_3 <- lm(mpg ~ poly(horsepower, 3), data = Auto, subset = train)
```

<br> 

Create predictions based on the model and test dataset (cross validate)

```{r}
pred_2 <- predict(model_2, Auto)[-train]
pred_3 <- predict(model_3, Auto)[-train]
```

<br>

Calculate MSE's for model_2 and model_3

```{r}
mean((Auto$mpg[-train] - pred_2)^2)
mean((Auto$mpg[-train] - pred_3)^2)
```

<br>

Seems that we've reduced the MSE from 26 to 19.8 by creating a quadratic fit. The cubic fit does very little to decrease the MSE. 

<br>

## k-fold cross validation

<br>

use cv.glm() function

```{r}
# has cv.glm() function
library('boot')

# typical model
glm_fit <- glm(mpg ~ horsepower, data = Auto)

# perform the cross validation
cv_err <- cv.glm(Auto, glm_fit)

# cross validation results
cv_err$delta
```

<br> 

Automating polynomial fits
```{r}
cv_error <- rep(0,5)

for (i in 1:5) {
  glm_fit <- glm(mpg ~ poly(horsepower, i), data = Auto)
  cv_error[i] <- cv.glm(Auto, glm_fit)$delta[1]
}

cv_error

plot(cv_error, type = 'b')
```

<br> 

No clear improvement from using higher order polynomials. 

<br>




